---
title: "Connective_Game_Analysis"
author: "Masoud Jasbi"
date: "August 18, 2016"
output: html_document
---
  
```{r packages, include=FALSE}
require(tidyverse)
require(ggplot2)
library(forcats)
library(rstan)
library(ggmcmc)

## for bootstrapping 95% confidence intervals
theta <- function(x,xdata) {mean(xdata[x])}
ci.low <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
ci.high <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.975)}

## reading the processed data
data_main <- read.csv("../2_processed_data/cardgame_data_processed.csv")

## isolating the trials with 3 values as DV
data_main_dv3 <- data_main %>%
  filter(dv_type==3) 
View(data_main_dv3)
```

```{r }
df_stan <- as_tibble(data_main_dv3) %>%
  filter(age_category == "Children") %>%
  select(-age_category, -dv_type) %>%
  mutate(response = fct_relevel(response, "Wrong","Kinda Right"), 
         age_centered_years = as.numeric(scale(age_year, scale=FALSE))) %>%
  filter(!is.na(response), 
         trial_type %in% c("X.XorY","X.XandY","XY.XorY","XY.XandY")) %>%
  mutate(sid = factor(sid), 
         trial_type = fct_relevel(factor(trial_type), "XY.XandY"))

df_stan_small <- df_stan %>%
  filter(sid %in% unique(data_main_dv3$sid)[1:5]) %>%
  mutate(sid = factor(sid))
```

```{r chisq-test}
criticals <- 
  data_main_dv3 %>% filter(!is.na(response), 
         trial_type %in% c("X.XorY","X.XandY","XY.XorY","XY.XandY"))
criticals$trial_type <- factor(criticals$trial_type)

tbl <- table(criticals$response, criticals$trial_type)
View(tbl)
chimodel <- chisq.test(tbl)

summary(chimodel)
```

```{r stats}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

cardgame_model <- "
data {
  int K; // number of responses
  int<lower=1> N; // number of trials
  int<lower=1> J; // number of subs
  int<lower=1, upper=3> response[N];
  int<lower=1, upper=J> subj[N];
  int x_and[N]; // one-animal conjunction. two-animals-conjunction is the intercept
  int x_or[N]; // one-animal disjunction
  int xy_or[N]; // two-animals disjunction
}

parameters {
  vector[J] sub_intercept; //subject intercepts
  real alpha; // 
  real beta_x_and; // one-animal 
  real beta_x_or; // 
  real beta_xy_or; // 
  ordered[K-1] c; // cutpoints
}

model {
  alpha ~ normal(0, 10);
  beta_x_and ~ normal(0, 10);
  beta_x_or ~ normal(0, 10);
  beta_xy_or ~ normal(0, 10);

  c ~ normal(0, 1);

  for (j in 1:J)
    sub_intercept[j] ~ normal(0, 10);

  for (n in 1:N)
    response[n] ~ ordered_logistic(alpha + 
        beta_x_and * x_and[n] + beta_x_or * x_or[n] + beta_xy_or * xy_or[n] + sub_intercept[subj[n]], c);
}
"

stan_data <- list(subj = as.integer(df_stan$sid),
                  response = as.integer(df_stan$response),
                  x_and = as.integer(df_stan$trial_type == "X.XandY"),
                  x_or = as.integer(df_stan$trial_type == "X.XorY"),
                  xy_or = as.integer(df_stan$trial_type == "XY.XorY"),
                  N = nrow(df_stan),
                  J = length(unique(df_stan$sid)),
                  K = 3)

cardgame_model <- stan(model_code = cardgame_model,
                       data = stan_data, 
                       warmup = 1000,
                       iter = 2000, chains = 4)
```

Extract the betas:

```{r}
cardgame_samples <- ggs(cardgame_model) 

fixefs <- filter(cardgame_samples, 
                 Parameter %in% c("alpha","beta_x_or",
                                  "beta_x_and","beta_xy_or"))

fixefs %>%  
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .025),
            hdp_upper = quantile(value, .975)) %>%
  knitr::kable(digits = 2)


fixefs %>%
  spread(Parameter, value) %>%
  mutate(and_mean = alpha - beta_x_and,
         x_mean = beta_x_and - beta_x_or, 
         or_mean = beta_x_or - beta_xy_or) %>%
  select(-alpha, -beta_x_and, -beta_x_or) %>%
  gather(Parameter, value, and_mean, x_mean, or_mean, beta_xy_or) %>%
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .025),
            hdp_upper = quantile(value, .975)) %>%
  knitr::kable(digits = 2)

```

Does intercept trade off with cutpoints?

```{r}
ggmcmc:ggs_pairs(cardgame_samples)

```

```


%>%
  mutate(x_and_or = beta_x_or-beta_x_and,
         xy_x_or = beta_xy_or - beta_x_or) %>% select(x_and_or, xy_x_or) %>% summarize(x_and_or_low = quantile(x_and_or,.025), 
            x_and_or_high = quantile(x_and_or,.975),
            xy_x_or_low = quantile(xy_x_or,.025), 
            xy_x_or_high = quantile(xy_x_or,.975)) %>% View()
```

```{r}
df_stan %>%
  group_by(trial_type, response) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n))
```

Explore chains

```{r}

ggmcmc(filter(cardgame_samples, !str_detect(as.character(Parameter), "sub_intercept")))
```

Old age model. 

```{r}
cardgame_model <- "
data {
  int K; // number of responses
  int<lower=1> N; // number of trials
  int<lower=1> J; // number of subs
  real age_year[N];
  int<lower=1, upper=3> response[N];
  int<lower=1, upper=J> subj[N];
  int x_and[N]; // xy and is the intercept
  int x_or[N];
  int xy_or[N];
}

parameters {
  // vector[3] beta; // fixed intercept and slopes
  // vector[J] u; //subject intercepts
  // real<lower=0> sigma_e; //error sd
  // real<lower=0> sigma_u; //subj sd
  real alpha;
  real beta_age;
  real beta_x_and;
  real beta_x_or;
  real beta_xy_or;
  real beta_age_x_and;
  real beta_age_x_or;
  real beta_age_xy_or;
  ordered[K-1] c; // cutpoints
}

model {
  alpha ~ normal(0, 10);
  beta_age ~ normal(0, 10);
  beta_x_and ~ normal(0, 10);
  beta_x_or ~ normal(0, 10);
  beta_xy_or ~ normal(0, 10);
  beta_age_x_and ~ normal(0, 10);
  beta_age_x_or ~ normal(0, 10);
  beta_age_xy_or ~ normal(0, 10);

  c ~ normal(0, 10);

  for (n in 1:N)
    response[n] ~ ordered_logistic(alpha + beta_age * age_year[n] + 
        beta_x_and * x_and[n] + beta_x_or * x_or[n] + beta_xy_or * xy_or[n] + 
        beta_age_x_and * age_year[n] * x_and[n] + 
        beta_age_x_or * age_year[n] * x_or[n] + 
        beta_age_xy_or * age_year[n] * xy_or[n], c);
}
"
```

To test this we look at the difference in paramter estimate iteration by iteration, the 95% confidence intervals for each iteration. 95% confidence interval on the difference between the two.
Report the means for the three betas with the confidence intervals and then a table for the differences with their confidence intervals
'
